---
title: "Searching for R Packages"
output: html_notebook
---

One of the most vexing problems for R users, and experienced alike, is to find packages with functions that can halp you do some new task. Recently, within the past year or so, a few packages have been written to help. The following recreates my strategy, more or less, to find what's out there for plotting [ROC curves](https://rviews.rstudio.com/2019/01/17/roc-curves/).

### Packages for Searching

There are two promary packages for searching `CRAN`: [packagefinder](https://CRAN.R-project.org/package=packagefinder) and [pkgesearch](https://CRAN.R-project.org/package=pkgsearch).

I will be using Gábor Csárdi's `pkgsearch` package, which is newer and has better documantation. You can find a blog post which used `packagefinder` [here](https://rviews.rstudio.com/2018/10/22/searching-for-r-packages/).


```{r, warning=FALSE, message=FALSE}

library(tidyverse)  # for data manipulation
library(dlstats)    # for package download stats
#library(packagefinder) # for searching packages
library(pkgsearch)  # for searching packages
```  

### Searching with `pkgsearch`
First we try some exploratory work to settle on a query. The documentation says:

Ranking:   

>The most important feature of a search engine is the ranking of the results. The best results should be listed first. pkgsearch uses weighted scoring, where a match in the package title gets a higher score than a match in the package desciption. It also uses the number of reverse dependencies and the number of downloads to weight the scores:

Preferring Phrases: 

>The search engine prefers matching whole phrases over single words. E.g. the search phrase “permutation test” will rank coin higher than testthat, even though testthat is a much better result for the single word “test”. 


```{r}
r1 <- as_tibble(pkg_search("Receiver Operating Characteristic"))
names(r1)

```

```{r}
r1 %>% select(score, package, revdeps, downloads_last_month)
```


```{r}
r2 <- as_tibble(pkg_search("ROC"))
r2 %>% select(score, package, revdeps, downloads_last_month)
```


```{r}
r3 <- as_tibble(pkg_search(" Receiver Operating Characteristic Curve, ROC"))
r3 %>% select(score, package, revdeps, downloads_last_month)
```

Let's cast a wide net and get the top 200 packages with the query: "Receiver Operating Characteristic Curve, ROC"

```{r}
rocPkgs  <- as_tibble(pkg_search(query="Receiver Operating Characteristic, ROC",size=400))

```


```{r}
topRoc <- rocPkgs %>% filter(score > 15)
dim(topRoc)
topRoc %>% select(score, package, title) %>% head
```

### The Tough Part
Looking through the packages selected with scores > 15, I selected 6 packages that I thought would meet most peoples plotting needs. I did not select any of the packages made to used with survial data. That would be a much more specialized need,


```{r}
shortList <- rocPkgs %>% filter(package %in% c("pROC","precrec","ROCit", "PRROC","ROCR","plotROC")) %>%
            select(score, package, revdeps, title, description )
shortList
```


### A Closer Look at Downloads

Now, we us Guangchuang Yu's `dlstats` package to take a closer looka at downloads.


```{r,message=FALSE,warning=FALSE}
downloads <- cran_stats(shortList$package)
topDown <- downloads %>% group_by(package) %>% summarize(downloads = sum(downloads)) %>%
           arrange(desc(downloads))

rocFinal <- right_join(shortList,topDown)
dim(rocFinal)
```   

```{r}
rocFinal %>% select(package, score, downloads)
```

```{r}

ggplot(downloads, aes(end, downloads, group=package, color=package)) +
  geom_line() + geom_point(aes(shape=package))

```




Note, that on February 8th the following query turned up 418 packages.
```{r,eval=FALSE}
allROC <- pkg_search(query="Receiver Operating Characteristic, ROC",
             format = "short",
              size=10000)
dim(allROC)
```









